---
title: "Coursera Practical Machine Learning - Prediction Assignment Writeup"
author: "Maheshwari"
date: "Friday, September 25, 2015"
output: html_document
---

#Summary


##Preparing
Loading all the required packages
```{r echo=FALSE}
    library(caret)
    library(rattle)
    library(rpart)
    library(rpart.plot)
    library(randomForest)
```

Loading the data

```{r, echo=TRUE}
    wholedata <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
    testingSet<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))
```

Cleaning the data 
Removed the rows of data which has "NA" and also the variables which are not required
or useful for the predictions from both the training set and the testing set.

```{r, echo=TRUE}
    NA_Count = sapply(1:dim(wholedata)[2],function(x)sum(is.na(wholedata[,x])))
    NA_list = which(NA_Count>0)
    
    wholedata = wholedata[,-NA_list]
    wholedata = wholedata[,-c(1:7)]
    wholedata$classe = factor(wholedata$classe)


    NA_Count1 = sapply(1:dim(testingSet)[2],function(x)sum(is.na(testingSet[,x])))
    NA_list1 = which(NA_Count1>0)
    testingSet = testingSet[,-NA_list]
    testingSet = testingSet[,-c(1:7)]
    dim(wholedata)
    dim(testingSet)
    #inTrain = createDataPartition(y=wholedata$classe,p=0.6,list=FALSE)
inTrain=createDataPartition(y=wholedata$classe, p=0.6, list=FALSE)
```

Creating the traning and testing data partitions

```{r, echo=TRUE}

    training <-wholedata[inTrain,]
    testing <- wholedata[-inTrain,]
```
## Tree Method
Using the Tree method to do the prediction of 'classe'.

```{r, echo=TRUE}

    modfit1 <- train(classe ~ .,method='rpart',data=training)
    fancyRpartPlot(modfit1$finalModel) 
    pred=predict(modfit1,newdata=testing)
    z=confusionMatrix(pred,testing$classe)
    z$table
    z$overall[1]
```
From the confusion matrix it is clear the accuracy of "0.49" for this model fit clearly shows "no purity" hence this model fit is rejected.

## Random Forest Method
Using Random forest method to do the prediction.
```{r, echo=TRUE}
    modfit2=randomForest(classe~., data=training, method='class')
    pred2 = predict(modfit2,testing,type='class') 
    z2=confusionMatrix(pred2,testing$classe)
    z2$table
```
Random forest prediction model provides 99% accurancy hence this model has been choosen to do predict the testing data set.
```{r, echo=TRUE}
    z2$overall[1]
```

#Conclusion
From the above results the random forest method provides the best fit model and it is been considered for testing the test data set to submit results.

```{r, echo=TRUE}
    pred3 =  predict(modfit2,testingSet,type='class')
    nofiles = length(pred3)
    for (i in 1:nofiles){
        filename =  paste0("problem_id",i,".txt")
        write.table(pred3[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
    pred3
```